NMRlipids IV: Headgroup \& glycerol backbone structures, and cation binding in bilayers with PS lipids

### 2018-06-07 Manuscript/manuscriptPS.tex
Update text

\todo{Details of the used spectrometer and maybe some other details should be given.}
\todo{Sample preparation should be described.}

@@

Figure \ref{PShgSIGNSsimpson} summarizes the experimental NMR results for POPS bilayer sample.
showing one single
    splitting for beta (which gives an order parameter equal to 0.12), and for alpha a superposition
    of a large splitting (order parameter equal to 0.09) and a very small splitting which cannot
    resolved with the available resolution.
    S-DROSS slice for beta clearly shows that the order parameter is negative,
    which is confirmed by SIMPSON simulations using the 
    The S-DROSS slice for alpha suggests that the higher order parameter is positive
    and the deviation towards negative values in the longer t1 times suggests
    that the smaller order parameter is negative.
    This is confirmed by SIMPSON simulation using value of 0.09
    for the larger alpha order parameter and the value of -0.02 for
    smaller (black curve). The value for the smaller
    alpha order parameter for SIMPSON calculation was taken from Fig 3 in Ref. \citenum{roux91},
    because resolution in 13C NMR experiments was nor high enough to determine
    numerical value for this. The S-DROSS curve from SIMPSON simulation with positive value
    for the smaller order parameter (dashed grey), did not agree with experiments 
    confirming the interpretation that the smaller order parameter is negative.
\begin{figure}[!htb]
  \centering
  \includegraphics[width=8.5cm]{../scratch/figIDEA.eps}
  \caption{\label{PShgSIGNSsimpson}
    (a) The headgroup region of the INEPT spectrum with headgroup and glycerol backbone carbons assigned.
    (b) 2D R-PDLF spectra for headgroup and glycerol backbone regions.
    (c) Slices for $\alpha$ and $\beta$ barbons.
    (d) Experimental SDROSS data (points) and SIMPSON simulations (lines).
    Order parameter values of -0.12 for the $\beta$-carbon, and 0.09 and -0.02
    for the larger and smaller $\alpha$-carbon slittings were used in the
    SIMPSON calculations. The S-DROSS curve from SIMPSON simulation with positive value
    for the smaller order parameter (dashed grey).
  }
  \todo{This is preliminary figure, should be polished.}
  \todo{Should we show slices for all the analyzed carbons in (c)?}
\end{figure}

@@

The headgroup and glycerol backbone order parameters of 
POPS measured in this work are compared to the literature values of
DOPS \cite{browning80} and POPC \cite{ferreira13} in Fig. \ref{HGorderParameters}. 
Our results for POPS are in good agreement with the previously reported
values for DOPS measured with $^2$H NMR. Significant differences are
observed between PC and PS lipids, especially at the headgroup region.  
Previous discussions in the literature have concluded that  
glycerol backbone structure is largely similar in
PC, PE, PG and PS lipids \cite{gally81}. The headgroup region was
found to be similar in PC, PE and PG lipid (assuming that the signs of PE and PG
order parameters are the same as in PC), while the PS headgroup was suggested
to be more rigid \cite{wohlgemuth80,buldt81}. 
The detailed structural differences between the headgroups is, however, not known.
\begin{figure}[]
  \centering
  \includegraphics[width=9.0cm]{../Figs/PCPScomp.pdf}
  \caption{\label{HGorderParameters}
    Headgroup and glycerol backbone order parameters of POPS measured in this work compared
    with values for DOPS ($^2$H NMR, 0.1M of NaCl) \cite{browning80} and 
    POPC  ($^{13}$C NMR) \cite{ferreira13} from literature. Signs for PS order parameters
    as measured in this work and signs for PC as measured in Refs \cite{??,ferreira16}.
    %POPG from \cite{borle85} contains 10nM PIPES, DPPG from \cite{wohlgemuth80} contains  10mM PIPES and 100mM NaCl,
    %DPPE from \cite{seelig76}, E.coliPE and E.coliPG are from \cite{gally81}.
  }
\end{figure}

>>>>

\todo{Details of the used spectrometer and maybe some other details should be given.} \\
\todo{Sample preparation should be described.} \\
\todo{How is the peak assignment done?}

@@

The INEPT and 2D R-PDLF experiments from POPS sample give well resolved spectras for all the
carbons in headgroup and glycerol backbone region, except for g$_3$ for which the resolution
was not sufficient to determine the numerical value of the order paramater (Fig. \ref{PShgSIGNSsimpson}).
Slices of the R-PDFL spectra (Fig. \ref{PShgSIGNSsimpson} C)) 
show a single splitting for the $\beta$-carbon with the order parameter value of 0.12,
and a superposition of a large and a very small splitting for the $\alpha$-carbon.
The larger splitting gives a order parameter value of 0.09, while the numerical value
from the small splitting cannot resolved with the available resolution.
Since only the absolute values of the PS headgroup order parameters were measured
previously \cite{browning80,roux91}, we used the S-DROSS experiment \cite{??} to determine the signs of
the order parameters. The S-DROSS slice for the $\beta$-carbon (Fig. \ref{PShgSIGNSsimpson} D)) clearly shows that the
order parameter is negative, which is confirmed by SIMPSON simulations.
The beginning of the S-DROSS slice suggests that the higher order parameter of
the $\alpha$-carbon is positive and the deviation towards negative values with the longer T$_1$ times suggests
that the smaller order parameter is negative. This is confirmed by a SIMPSON simulation
where the value of -0.02 was taken from $^2$H NMR experiment \cite{roux91} for the smaller order parameter.
The literature value was used because the
resolution of our experiment was not sufficient to determine the
small value of the order parameter. The S-DROSS curve from
SIMPSON simulation with a positive value for the smaller order parameter
(dashed grey in Fig. \ref{PShgSIGNSsimpson} D)) did not agree with the experiment, 
confirming the interpretation that the smaller order parameter is negative.
\begin{figure}[!htb]
  \centering
  \includegraphics[width=8.5cm]{../scratch/figIDEA.eps}
  \caption{\label{PShgSIGNSsimpson}
    (a) The headgroup region of the INEPT spectrum with headgroup and glycerol backbone carbons assigned.
    (b) 2D R-PDLF spectra for headgroup and glycerol backbone regions.
    (c) Slices for $\alpha$ and $\beta$ barbons.
    (d) Experimental SDROSS data (points) and SIMPSON simulations (lines).
    Order parameter values of -0.12 for the $\beta$-carbon, and 0.09 and -0.02
    for the larger and smaller $\alpha$-carbon slittings were used in the
    SIMPSON calculations. The S-DROSS curve from SIMPSON simulation with positive value
    for the smaller order parameter (dashed grey).
  }
  \todo{This is preliminary figure, should be polished.}
  \todo{Should we show slices for all the analyzed carbons in (c)?}
\end{figure}

@@

The headgroup and glycerol backbone order parameters of 
POPS measured in this work are in good agreement with the previously reported
values from $^2$H NMR experiments of DOPS \cite{browning80} (Fig. \ref{HGorderParameters}).
The $\beta$-carbon order parameter is significantly more negative and $\alpha$-carbon
experiences a significant forking in PS headgroup when compared with
the values previously measured for POPC \cite{ferreira13} (Fig. \ref{HGorderParameters}).  
These features have been intepreted to arise from a rigid PS headgroup
conformation, stabilized by hydrogen bonds or electrostatic
interactions \cite{browning80,buldt81}, but detailed structrural interpretation is not
available. 
%The detailed structural differences between the headgroups is, however, not known.
\begin{figure}[]
  \centering
  \includegraphics[width=9.0cm]{../Figs/PCPScomp.pdf}
  \caption{\label{HGorderParameters}
    Headgroup and glycerol backbone order parameters of POPS measured in this work compared
    with values for DOPS ($^2$H NMR, 0.1M of NaCl) \cite{browning80} and 
    POPC  ($^{13}$C NMR) \cite{ferreira13} from literature. Signs for PS order parameters
    as measured in this work and signs for PC as measured in Refs \cite{??,ferreira16}.
    %POPG from \cite{borle85} contains 10nM PIPES, DPPG from \cite{wohlgemuth80} contains  10mM PIPES and 100mM NaCl,
    %DPPE from \cite{seelig76}, E.coliPE and E.coliPG are from \cite{gally81}.
  }
  \todo{There should be values in \cite{roux90} which should be added.}
\end{figure}
